---
title: "Gov 50 Final Project"
author: "Anna MacLennan"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to the 2022 midterm elections and Asian American voters, or other data related to Asian Americans and American politics. I could potentially study Asian American voting behavior over time related to some other factor such as specific ethnicity, what generation immigrant they are, etc. Or I could look at what issues they rate as being most important to them, or whether thay're more likely to vote for Asian candidates regardless of party affiliation.

```{r}
knitr::opts_chunk$set(echo = FALSE)



```
```{r}

library(tidyverse)
library(ggplot2)
#Options
options(
  scipen = 9,
  width = 100)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center")


aapi <- read.csv("aapi-final-project-data.csv")
```


```{r}
#creating variable for whether someone has experienced discrimination
#is a 1 if respondent indicated they had experience emotional bullying,
#disrespect/being ignored, and/or physical attack
aapi <- aapi |>
  drop_na(emo_bully, disrespect_ignored, phys_attack, racism_issue) |>
  mutate(
    discrim = if_else(emo_bully == 0 & disrespect_ignored == 0 & phys_attack == 0, 0, 1)
  )

filtered_aapi <- aapi |>
  filter(orig_region == "s" | orig_region == "se" | orig_region == "e") |>
  mutate(
    orig_north_south = case_when(
      orig_region == "s" ~ 0,
      orig_region == "e" ~ 1,
      orig_region == "se" ~ 0,
    ))

#Does experiencing discrimination make you care more about racism?
aapi |>
  mutate(discrim = if_else(discrim == 0, "No Discrim", "Yes Discrim")) |>
  group_by(discrim) |>
  summarize(avg_careaboutracism = mean(racism_issue), n = n()) |>
  pivot_wider(names_from = discrim, values_from = avg_careaboutracism)

#Does your region of origin affect whether you care about racism?
filtered_aapi |>
  #filter(orig_country == "China" | orig_country == "India") |>
  group_by(orig_region) |>
  summarize(avg_careaboutracism = mean(racism_issue), n = n()) 
```

```{r}
racism_affected <- lm(racism_issue ~ discrim + orig_region, data = filtered_aapi)

racism_affected
```

